---
title: "Summer Olympics Analysis"
author: "MJ"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
---

## Intro

Below analysis is based on [Kaggle: Olympic Sports and Medals](https://www.kaggle.com/the-guardian/olympic-games/data) dataset, which aggregates all of the medals granted to Olympians durring Summer Games.

```{r, dane, echo=FALSE}
oly <- read.csv(
  file="summer.csv",
  header = TRUE, sep=";",
  dec = ".")
```
### Basic information

Dataset consists of `r dim(oly)[2]` variables and `r nrow(oly)` observations.

1. **Year**
2. **City**
3. **Sport**
4. **Discipline**
5. **Athlete**
6. **Country**
7. **Gender**
8. **Event**
9. **Medal**

## Tables and plots

### **List of the best scoring (number of medals) countries in history**

```{r, tab1, echo=FALSE, warning=FALSE,message=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
ctry1<- data.frame(head(sort(table(oly$Country), decreasing = TRUE), 10))
ctry2<-data.frame(count(oly, Country, Medal))
mrgd<-ctry2[ctry2[,1] %in% ctry1[,1],]

colnames(ctry1) <- c("Country", "Number_of_medals")
kable(ctry1)


ggplot(data = mrgd, mapping = aes(x = mrgd[,1], y=mrgd[,3])) +
  geom_bar(stat="identity",aes(fill=Medal))+scale_fill_manual(values=c("sienna","Gold", "Grey"))+
  theme_bw()+
  labs(y="Number of medals", 
       x="Country")

```

### **Medals for Poland in chosen sports**

```{r, tab2, echo=FALSE,warning=FALSE}
pol<- oly[ which(oly$Country=='POL'),  ]
pol2<-data.frame(count(pol, Sport))

colnames(pol2) <- c("Sport", "Number_of_medals")
kable(pol2)

ggplot(data = pol2, mapping = aes(x = Sport, y=Number_of_medals)) +
  geom_point(stat="identity", col="Red")+
  theme_bw()+
  labs(y="Number of medals", 
       x="Discipline")+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### **Medals in Olympic Weightlifting**

```{r, tab3, echo=FALSE,warning=FALSE}
wl0<- data.frame(oly[ which(oly$Sport=='Weightlifting'),  ])
wl1<-data.frame(count(wl0, Country))

colnames(wl1) <- c("Country", "Number_of_medals")
kable(wl1)

ggplot(data = wl1, mapping = aes(x = Country, y=Number_of_medals)) +
  geom_bar(stat="identity", fill="blue")+
  theme_bw()+
  labs(y="Number of medals", 
       x="Country")+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
