---
title: "Titanic - Simple Analysis of Survival"
author: "MJ"
output:
  html_document
---

## Intro

`Titanic` dataset is built-in base *R*. It includes information about passengers and their survival from the catastrophe. More detailed dataset can be found here: [Kaggle: Titanic Competition](https://www.kaggle.com/c/titanic).

### Variables
```{r Titanic, echo=FALSE}
podsum<-summary(Titanic)
```

Dataset has `r podsum[1]` factors and `r podsum[2]` cases.

1. **Age**:
  * Adults
  * Children

2. **Sex**:
  * Women
  * Men

3. **Travel Class**:
  * 1st Class
  * 2nd Class
  * 3rd Class
  * Staff

4. **Survived**
  * Yes
  * No


### Data

Aggregated data was presented in below tables.

**Survival dependent on Travel Class, Age, Sex:**

```{r tab3, echo=FALSE, warning=FALSE}
library(knitr)
titanicdf<-data.frame(Titanic)
class <-tapply(titanicdf$Freq, titanicdf[,c(1,4)], FUN = sum)
sex <-tapply(titanicdf$Freq, titanicdf[,c(2,4)], FUN = sum)
age <-tapply(titanicdf$Freq, titanicdf[,c(3,4)], FUN = sum)
merged <- rbind(class, sex, age)
kable(merged)

```

### Plots

```{r wyk1, echo=FALSE}
op <- par(mfrow=c(1,3))
barplot(class, 
        main="Survival by passenger status",
        xlab="Survived",
        ylab="Number of passengers",
        legend =row.names(class),
        col=c("darkgreen","green", "lightgreen", "blue"),
        beside=T
        )


barplot(sex, 
        main="Survival by sex",
        xlab="Survived",
        ylab="Number of passengers",
        legend =row.names(sex),
        col=c("darkgreen","green"),
        beside=T
        )


barplot(age, 
        main="Survival by age",
        xlab="Survived",
        ylab="Number of passengers",
        legend =row.names(age),
        col=c("darkgreen","green"),
        beside=T
        )

par(op)
```
